One-dimensional turbulence (ODT): computationally efficient modeling and simulation of turbulent flows

### 2020-05-23 manuscript/manuscript.tex
Updated text of model description section. Working on software architecture section. Flow chart in progress elsewhere.

The one-dimensional turbulence model (ODT) functionally reverses the LES approach, modeling large-scale turbulent advection and directly resolving small-scale flow structures, simulating the full range of length and time scales in a single dimension. Because large-scale structures are much easier to study and model than small-scale structures, ODT mitigates or sidesteps many of the subgrid modeling issues that complicate LES. Previous studies show that ODT can attain accuracy comparable to DNS at a fraction of the computational cost \cite{Lignell_2015,Abboud_2015}, making it an attractive tool for simulating turbulent flows. Because the model is one-dimensional, it is limited to homogeneous or boundary layer flows such as jets, wakes, and mixing layers; such flows, however, are extremely common in both nature and turbulence research. ODT's computational efficiency and resolution of a full range of scales make it a valuable tool that complements experimental studies and other simulation tools like DNS and LES.

@@

The ODT model is described in detail in the literature \cite{Kerstein_1999,Kerstein_2001,Ashurst_2005,Lignell_2018,Lignell_2013}; only a brief explanation will be given here. In ODT, turbulent advection is modeled with stochastic processes called eddy events, which punctuate the solution of unsteady, one-dimensional transport equations for mass, momentum, and enthalpy. The ODT code uses a Lagrangian finite-volume formulation for diffusive advancement that includes adaptive mesh refinement \cite{Lignell_2013}. In this approach, mass remains constant inside each grid cell while cell volumes increase or decrease according to dilation. Because the ODT model is one-dimensional, it is limited to homogeneous or boundary-layer flows, such as jets, wakes, and mixing layers; these types of flows, however, are common in nature and central to turbulence research.

Eddy events occur as a Poisson process in accordance with their eddy rates, where a given eddy event of size $l$ and location $x_0$ has an eddy timescale $t$ and an associated eddy rate $1/t$. Three user-defined ODT parameters control the eddy event process: the eddy rate parameter $C$ scales the rate of occurrence of the eddies; the viscous penalty parameter $Z$ suppresses small eddies; and the large eddy suppression parameter $\beta$ constrains eddies such that they do not reach over the elapsed simulation time.

@@

@@

The ODT code is a relatively self-contained C++ package. The system of nonlinear ODEs is solved using CVODE \cite{Hindmarsh_2020} and user input files are processed with YAML \cite{Beder_2008}, both of which are installed locally during the ODT build process. For reacting flow cases, chemical kinetics and transport are handled by Cantera \cite{Goodwin_2018}, which must be previously installed by the user.

@@

>>>>

The one-dimensional turbulence model (ODT) functionally reverses the LES approach, modeling large-scale turbulent advection and directly resolving small-scale flow structures, simulating the full range of length and time scales in a single dimension. Because large-scale structures are much easier to study and model than small-scale structures, ODT mitigates or sidesteps many of the subgrid modeling issues that complicate LES. Previous studies show that ODT can attain accuracy comparable to DNS at a fraction of the computational cost \cite{Lignell_2015,Abboud_2015}, making it an attractive tool for simulating turbulent flows. Because the ODT model is one-dimensional, it is limited to homogeneous or boundary-layer flows, such as jets, wakes, and mixing layers; these types of flows, however, are common in nature and central to turbulence research. ODT's computational efficiency and resolution of a full range of scales make it a valuable tool that complements experimental studies and other simulation tools like DNS and LES.

@@

The ODT model is described in detail in the literature \cite{Kerstein_1999,Kerstein_2001,Ashurst_2005,Lignell_2018,Lignell_2013}; only a brief explanation will be given here. In ODT, turbulent advection is modeled with stochastic processes called eddy events, which punctuate the solution of unsteady, one-dimensional transport equations for mass, momentum, and enthalpy. The ODT code uses a Lagrangian finite-volume formulation for diffusive advancement in which mass stays constant within each grid cell while cell volumes increase or decrease according to cell dilation via an adaptive mesh refinement \cite{Lignell_2013}.

Transport equations for mass, momentum, and enthalpy in the temporal formulation of ODT take the following generic form, derived from the Reynolds Transport Theorem \cite{Cengel_2010} for a given scalar quantity per unit mass $\beta$: 
\begin{equation} 
\label{eq:transport}
	\frac{\mathrm{d}\beta}{\mathrm{d}t} = -\frac{j_{\beta,e}A_{x,e}-j_{\beta,w}A_{x,w}}{\rho V} + \frac{S_{\beta}}{\rho V}.
\end{equation}
Here, $j_{\beta}$ is the diffusion flux of scalar $\beta$ across the cell face area $A_x$ where the subscripts $e$ and $w$ refer to the "east" and "west" faces of the grid cell, respectively. $S_{\beta}$ is the Lagrangian source term derived from the conservation law for $\beta$, $\rho$ represents mass density, and $V$ represents cell volume. In practice, we refer to the left hand term on the right side of Equation~\ref{eq:transport} as the "mixing term" and the right hand term on the right side of Equation~\ref{eq:transport} as the "source term". The generic transport equation differs slightly in the spatial formulation of ODT, but its form is the same, so we omit it here for brevity. The system of ordinary differential equations (ODEs) that results is well behaved at all grid points and in all geometries in their finite-volume forms. For details on transport equation derivation and use in both the temporal and spatial formulations of ODT, see Lignell et al. \cite{Lignell_2018}.

Eddy events occur as a Poisson process in accordance with their eddy rates, where a given eddy event of size $l$ and location $x_0$ has an eddy timescale $t$ and an associated eddy rate $1/t$. Three user-defined ODT parameters control the eddy event process: the eddy rate parameter $C$ scales the rate of occurrence of the eddies; the viscous penalty parameter $Z$ suppresses small eddies; and the large eddy suppression parameter $\beta$ constrains eddies such that they do not reach over the elapsed simulation time. Sampled eddies that do not fit the defined parameters are rejected and not applied to the domain.

@@

Eddy events occur concurrently with diffusive advancement via solution of the system of unsteady one-dimensional transport equations. In this way, the ODT code marches in time or space until it reaches its end point. Due to the stochastic nature of eddy events, each ODT simulation, or realization, is different, even when it is provided with the same input parameters. In order to obtain statistically stable data for a given set of parameters, we run many realizations with the same input parameters and time-average them. This is done via post-processing tools, which are provided in the ODT package.

@@

The ODT package consists primarily of an object-oriented C++ code responsible for running flow simulation cases and generating data. The package also contains auxiliary data processing and visualization tools, written mostly in Python. Within the main download package, several directories organize the code:
\begin{itemize} \itemsep0em
	\item \texttt{source}: ODT source code
	\item \texttt{build}: compiles and builds source code via CMake
	\item \texttt{run}: ODT executable and run scripts
	\item \texttt{data}: data generated by source code
	\item \texttt{input}: input files used to specify case variables
	\item \texttt{post}: data post-processing and visualization scripts
	\item \texttt{doc}: code documentation generated by Doxygen
\end{itemize}

% source code stuff
In an individual simulation, variables live within in the \texttt{domain} class; wherever these variables are used within the code, pointers refer back to these locations. Key quantities are referred to as line variables, and their individual properties and functions exist within child classes of the \texttt{domainvariables} parent class.

The system of nonlinear ODEs is solved using CVODE \cite{Hindmarsh_2020} and user input files are processed with YAML \cite{Beder_2008}, both of which are installed locally during the ODT build process. For reacting flow cases, chemical kinetics and transport are handled by Cantera \cite{Goodwin_2018}, which must be previously installed by the user.

% You can talk about the basic structure: domain.cc is the main object that holds pointers to everything else. There are back pointers to the domain, so that everyone can see everyone else, as needed. The key quantities are line variables (arrays stored at each point). These are the domainvariables. The domain class has a vector of pointers to these, and also pointers to specific variables to access them by name. This enables both generic looping over variables to have each variable compute, for example, its source term, and also allow specific access to needed variables like the grid positions, etc. The variables are either transported or not. The variables needed for a given simulation are set up in the domaincases class. In addition, there are general parameters. You could discuss the main classes, or maybe show a class flow diagram. These can be easily generated, and doxygen also does some of this.

@@

%\subsection{General Workflow}
%\label{sub:workflow}
%
%Every ODT simulation shares common steps to a successful run. 
%\begin{enumerate}
%	\item Build and compile the code. 
%	\item Choose a case type and chemical mechanism (when required) and alter the appropriate input file(s). 
%	\item Within the \texttt{run} directory, alter the run script to be used with a case name and other required information.
%	\item Run the case from the \texttt{run} directory using the chosen run script.
%	\item Run the appropriate post-processing scripts, located in the \texttt{post} directory, on the raw data for interpretation and visualization.
%\end{enumerate}
